An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-09-23 nips_2016.tex
Modified Secdtion 6.2

\begin{figure}[t]
	\centering
	\includegraphics[width=\linewidth]{GaussianMixtureResult/first_second_part_gaussian.png}
	\caption{Value of first and second part of equation (9) in~\cite{icml2014c1_bardenet14} with respect to actual minibatch size}
	\label{fig:gauss_bardenet}
\end{figure}

@@

We next use logistic regression for the binary classification of 1s versus 7s in the MNIST
dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007 and 1000 training and testing
points, respectively (we used a random subset of the test data). The proposer is again a random walk
with covariance matrix $0.05I$ for the $784\times 784$ identity matrix $I$. We set the posterior
temperature at $T=1000$. We set the minibatch size $m=100$ and compare with ~\cite{cutting_mh_2014} with
tolerance 0.005 and with ~\cite{icml2014c1_bardenet14} with error bound control parameter $p = 2, \delta = 0.01$.

Figure~\ref{fig:logistic_fig} shows the prediction accuracy and log likelihood on the test set as a
function of the cumulative training data points processed. Our test increments the cumulative data
by a fixed amount per iteration, but the other two methods may require more data per iteration.  We
see that our minibatch MH test is more efficient; it has similar or better performance while
consuming fewer data points.

\begin{figure}[t]
    \centering
    \begin{subfigure}{.4\textwidth}
     \centering
     \includegraphics[width=1\linewidth]{LogisticRegressionResult/acc_logistic}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
     \centering
    \includegraphics[width=1\linewidth]{LogisticRegressionResult/ll_logistic}
\end{subfigure}
\caption{Logistic regression performance (accuracy/log likelihood) and minibatch size analysis}
\label{fig:logistic_fig}
\vspace{-10pt}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=1\linewidth]{LogisticRegressionResult/minibatch_size_logistic.png}
	\caption{Histograms of minibatch sizes of logistic regression parameter estimation}
	\label{fig:logistic_minibatch}
\end{figure}

%\begin{figure}[t]
  %  \centering
    %\includegraphics[width=1\linewidth]{CI_v02.png}
    %\caption{
   % Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
    %}
    %\label{fig:exp_3}
    %\vspace{-10pt}
%\end{figure}

The plot of Figure ~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes. Choosing an initial minibatch size of 100, 
the cut MH method, adaptive MH method achieves final average minibatch size of 585 and 2731 respectively, while our method achieves 
that of 100, showing significant better performance than the benchmark test methods.

%The third plot of Figure~\ref{fig:logistic_minibatch} shows performance based on minibatch size. The
%parameter is evaluated on the test set after training iterations 700, 1000, and 1500. For those
%three cases, when the minibatch size increases beyond about 10, the performance of the sampled data
%will not improve with a large minibatch size. This makes sense because the error of our MH test is
%proportional to our estimation error of ${\rm std}(\Delta')$, and as the minibatch increases, ${\rm
%std}(\Delta')$ decreases. Beyond a certain standard deviation (or minibatch) threshold, the error in
%our minibatch MH test is negligible.

%Similar as the Subsection~\ref{ssec:gaussians}, we use Figure~\ref{fig:exp_3}
%to reveal the accuracy of our estimation of ${\rm std}(\Delta')$ and stability
%of $\rho/\sigma^3$.

%% \subsection{Neural Network Optimization}\label{ssec:nets}
%% 
%% In this experiment, we apply our minibatch MH test with a Stochastic Gradient Hamiltonian
%% Monte Carlo (SGHMC)~\cite{sghmc_2014} proposer to sample the complete parameters for a fully
%% connected neural network. We use the Higgs data set from the UCI dataset~\cite{Lichman:2013}, which
%% is a binary classification task with one million instances, each of which are 28-dimensional. The
%% network we use has four layers: the first is the input, the last is the softmax, and the
%% intermediate ones apply the sigmoid activation unit.  We also employ dropout and batch
%% normalization~\cite{icml2015_ioffe15}.  In order to control ${\rm std}(\Delta')$, we initialize the
%% temperature at 1000, and adjust it at iteration $i$ according to $T_i = \max\{1,
%% 1000/(i+1)^{0.5}\}$.
%% 
%% For our comparison baseline, we train the same network architecture using the Adaptive Gradient
%% Descent method~\cite{adapGrad}. Since the frequencies of the features are quite different, we use
%% the same strategy as in Adaptive Gradient to rescale the gradient before applying SGHMC. We
%% implement our MH test and SGHMCin the open-source BIDMach project~\cite{canny2013bidmach}.  
%% 
%% Figure~\ref{fig:nnet_fig} illustrates our experiment results and reveals that SGHMC with our
%% minibatch MH test achieves higher log likelihood than the baseline. By comparing SGHMC with and
%% without our minibatch MH test, we see that our MH test has limited benefit. Because we decrease the
%% step size quickly, the acceptance rate of SGHMC is already high, so there is little need to test.
%% % Daniel's note: Haoyu said he cannot increase the step size without it violating the variance
%% % condition.
%% 
%% \begin{figure}[t]
%%     \centering
%%     \includegraphics[width=1\linewidth]{exp3}
%%     \caption{Log likelihood versus number of passes for optimization of neural network.}
%%     \label{fig:nnet_fig}
%%     \vspace{-10pt}
%% \end{figure}

>>>>

\begin{figure}[t]
	\centering
	\includegraphics[width=\linewidth]{GaussianMixtureResult/first_second_part_gaussian.png}
	\caption{
    Value of first and second part of equation (9)
    in~\cite{icml2014c1_bardenet14} with respect to actual minibatch size.
    {\color{blue} Daniel: probably debatable on whether we should include this.
    Let's talk it over.}
    }
	\label{fig:gauss_bardenet}
\end{figure}

@@

We next use logistic regression for the binary classification of 1s versus 7s in
the MNIST dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007
training and 1000 testing elements (we used a random subset of the test data).
For the proposer, we use a random walk with covariance matrix $0.05I$ for the
$784\times 784$ identity matrix $I$. We set the posterior temperature at
$T=1000$ and use the minibatch size $m=100$ and compare
with~\cite{cutting_mh_2014} with tolerance 0.005 and
with~\cite{icml2014c1_bardenet14} with error bound control parameters $p = 2$
and $\delta = 0.01$. {\color{blue} Daniel: just observed this, but why are we
using a subset of the test data? We really should use all 2000 (I think it's
2000, right?).}

Figure~\ref{fig:logistic_fig} shows the prediction accuracy and log likelihood
on the test set as a function of the cumulative training data points processed.
Our test increments the cumulative data by a fixed amount per iteration
{\color{blue} (Daniel: are we sure about ``fixed amount per iteration'' because
our Algorithm 1 implies we may have to increment the minibatch size, but it
happens \emph{less frequently} than the other methods.)}, but the other two
methods may require more data per iteration. We see that our minibatch MH test
is more efficient; it has similar or better performance while consuming fewer
data points.

\begin{figure}[t]
    \centering
    \begin{subfigure}{.4\textwidth}
     \centering
     \includegraphics[width=1\linewidth]{LogisticRegressionResult/acc_logistic}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
     \centering
    \includegraphics[width=1\linewidth]{LogisticRegressionResult/ll_logistic}
\end{subfigure}
\caption{
    Logistic regression performance (accuracy/log likelihood) and minibatch size
    analysis. {\color{blue} Daniel: let's use matplotlib's ``subplot''
    functionality, so that the images appear the same size. We can handle this
    later; it will be easy.}
    }
\label{fig:logistic_fig}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=1\linewidth]{LogisticRegressionResult/minibatch_size_logistic.png}
	\caption{
    Counts of minibatch sizes needed for the three methods in the MNIST
    classification task.
    }
	\label{fig:logistic_minibatch}
\end{figure}

Figure~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes for
all three methods. With an initial minibatch size of 100, Korattikara 2014 and
Bardenet 2014 achieve average minibatch size of 585 and 2731 respectively over
the MNIST classification task, while our method achieves that of 100, showing
significant better performance than the benchmark test methods. {\color{blue}
Daniel: these numbers don't match the figure titles.}
